---
title: "Reproducible Research: Peer Assessment 1"
output:
html_document:
keep_md: true
---

## Loading and preprocessing the data
  
Load the data (i.e. read.csv()).
  
```{r}
activity_set <- read.csv('activity.csv')
```

Process/transform the data (if necessary) into a format suitable for your analysis.

```{r}
activity_set$date <- as.Date(activity_set$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day.

```{r}
total_steps_by_day <- aggregate(activity_set$steps, 
                         by=list(activity_set$date), 
                         FUN=sum, na.rm=TRUE)
colnames(total_steps_by_day) <- c("date", "steps")
hist(total_steps_by_day$steps, col="red", 
     main="Daily Steps", xlab="Number of steps")
```

Calculate and report the mean and median total number of steps taken per day.

```{r}
mean <- mean(total_steps_by_day$steps)
median <- median(total_steps_by_day$steps)
``` 

The mean is `r mean` and the median is `r median`.

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
average_daily_activity <- aggregate(activity_set$steps, 
                                    by=list(activity_set$interval), 
                                    FUN=mean, na.rm=TRUE)
colnames(average_daily_activity) = c("interval", "steps")
plot(average_daily_activity$interval, 
     average_daily_activity$steps, col="red", type = "l", 
     main="Average daily steps", xlab="Interval", ylab="Number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval_with_maximum_number_of_steps <- 
  average_daily_activity$interval[which.max(average_daily_activity$steps)]
```

The maximum number of steps is in interval `r interval_with_maximum_number_of_steps`.

## Imputing missing values

## Are there differences in activity patterns between weekdays and weekends?
