---
title: "Reproducible Research: Peer Assessment 1"
output:
html_document:
keep_md: true
---

## Loading and preprocessing the data
  
Load the data (i.e. read.csv()).
  
```{r}
activity_set <- read.csv('activity.csv')
```

Process/transform the data (if necessary) into a format suitable for your analysis.

```{r}
activity_set$date <- as.Date(activity_set$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day.

```{r}
total_steps_by_day <- aggregate(activity_set$steps, 
                         by=list(activity_set$date), 
                         FUN=sum, na.rm=TRUE)
colnames(total_steps_by_day) <- c("date", "steps")
hist(total_steps_by_day$steps, col="red", 
     main="Daily Steps", xlab="Number of steps")
```

Calculate and report the mean and median total number of steps taken per day.

```{r}
mean <- mean(total_steps_by_day$steps)
median <- median(total_steps_by_day$steps)
``` 

The mean is `r mean` and the median is `r median`.

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
average_daily_activity <- aggregate(activity_set$steps, 
                                    by=list(activity_set$interval), 
                                    FUN=mean, na.rm=TRUE)
colnames(average_daily_activity) = c("interval", "steps")
plot(average_daily_activity$interval, 
     average_daily_activity$steps, col="red", type = "l", 
     main="Average daily steps", xlab="Interval", ylab="Number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval_with_maximum_number_of_steps <- 
  average_daily_activity$interval[which.max(average_daily_activity$steps)]
```

The maximum number of steps is in interval `r interval_with_maximum_number_of_steps`.

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).

```{r}
na_steps <- table(is.na(activity_set$steps))
```

There is `r na_steps[2]` rowns with NAs.

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
no_na_activity_set <- activity_set
no_na_activity_set$steps[is.na(no_na_activity_set$steps)] <- 
  average_daily_activity$steps
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
no_na_total_steps_by_day <- aggregate(no_na_activity_set$steps, 
                         by=list(no_na_activity_set$date), 
                         FUN=sum, na.rm=TRUE)
colnames(no_na_total_steps_by_day) <- c("date", "steps")
hist(no_na_total_steps_by_day$steps, col="blue", 
     main="Daily Steps", xlab="Number of steps")
hist(total_steps_by_day$steps, col="red", add=TRUE, 
     main="Daily Steps", xlab="Number of steps")
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```

```{r}
no_na_mean <- mean(no_na_total_steps_by_day$steps)
no_na_median <- median(no_na_total_steps_by_day$steps)
```

The mean is `r no_na_mean` and the median is `r no_na_median`.

## Are there differences in activity patterns between weekdays and weekends?
